---
title: "Text mining project"
author: "Krystian Andruszek, Ewelina Osowska and Ewa Sobolewska"
date: "13 01 2020"
output:
  rmarkdown::html_document: 
    theme: yeti
    toc: true
    toc_float: true
    toc_depth: 2
---


```{r, warning=FALSE, message=FALSE}
# Loading packages
library(readtext)
library(readxl)
library(dplyr)
library(tidyverse)
library(tidytext)
library(tm)
library(textstem)
library(wordcloud)
library(slam)
library(topicmodels)
library(SentimentAnalysis)
library(igraph)
library(ggraph)
library(widyr)
library(viridis)
library(plotly)
library(ggplot2)
library(lubridate)
library(wordcloud2)
library(stringr)
library(tidyr)
```

## Introduction

The main objective of this paper is to analyse the Federal Open Market Commitee statements using text mining methods and tools provided by R. We start with basic analysis of the length of each statement, through word counts, sentiment and correlation analysis and also topic modelling.

## What is Federal Open Market Committee?

Federal Open Market Committee (FOMC) is the body of the central bank of United States (the Federal Reserve System). Its main duties is setting the national monetary policy. The FOMC holds eight regularly scheduled meetings per year. At these meetings, the Committee reviews economic and financial conditions, determines the appropriate stance of monetary policy, and assesses the risks to its long-run goals of price stability and sustainable economic growth. The FOMC consists of 12 voting members: seven members of the Board of Governors, the president of the Federal Reserve Bank of New York and 4 of the remaining 11 Reserve Bank presidents, who serve one-year terms on a rotating basis. All 12 of the Reserve Bank presidents attend FOMC meetings and participate in FOMC discussions, but only the presidents who are Committee members at the time may vote on policy decisions. FOMC meetings typically are held eight times each year in Washington, D.C., and at other times as needed. 

## How are statements organized?

The Committee releases a public statement immediately after each FOMC meeting. Each statement follows very similar structure. Firstly, the general background of the economic situation is presented. Then the Commitee introduces the value of the established federal funds rate and also share predictions. At the end, there are listed names of people which voted for the FOMC monetary policy action.

## Data description

We sourced the data by scraping the statements from the Federal Reserve official website [^fomc] using Python. In the scraping algorithm we limited the content only to FOMC announcment, omitting the names of voters listed in the last paragraph. The analysed period includes years from 2006 to 2018 which resulted in obtaining 107 documents. 

```{r}
# Loading scrapped statements
DATA_DIR <- "~/Desktop/FOMC-text-mining/Statements"

fomc_2006 <- readtext(paste0(DATA_DIR, "/2006/*"))
fomc_2007 <- readtext(paste0(DATA_DIR, "/2007/*"))
fomc_2008 <- readtext(paste0(DATA_DIR, "/2008/*"))
fomc_2009 <- readtext(paste0(DATA_DIR, "/2009/*"))
fomc_2010 <- readtext(paste0(DATA_DIR, "/2010/*"))
fomc_2011 <- readtext(paste0(DATA_DIR, "/2011/*"))
fomc_2012 <- readtext(paste0(DATA_DIR, "/2012/*"))
fomc_2013 <- readtext(paste0(DATA_DIR, "/2013/*"))
fomc_2014 <- readtext(paste0(DATA_DIR, "/2014/*"))
fomc_2015 <- readtext(paste0(DATA_DIR, "/2015/*"))
fomc_2016 <- readtext(paste0(DATA_DIR, "/2016/*"))
fomc_2017 <- readtext(paste0(DATA_DIR, "/2017/*"))
fomc_2018 <- readtext(paste0(DATA_DIR, "/2018/*"))
# Binding data
statements <- rbind(fomc_2006, fomc_2007, fomc_2008, fomc_2009, fomc_2010, fomc_2011,
                    fomc_2012, fomc_2013, fomc_2014, fomc_2015, fomc_2016, fomc_2017, fomc_2018)
# Removing files from memory
remove(fomc_2006, fomc_2007, fomc_2008, fomc_2009, fomc_2010, fomc_2011,
       fomc_2012, fomc_2013, fomc_2014, fomc_2015, fomc_2016, fomc_2017, fomc_2018)
```

## Text preparation

We start our work on statments with the initial preprocessing of the dataset. It consists of two columns: doc_id and text. Doc_id is sourced from each statement's website link. Text is just a content of the statement. 

```{r}
head(statements, 1)
```

```{r}
# adding an unique ID
statements <- statements %>% mutate(ID = 1:n())
# setting column names 
colnames(statements) <- c("Date", "Text", "ID")
# modification of doc_id column - changing it to date column
statements$Date <- gsub(".txt", "", statements$Date)
statements$Date <- as.Date(statements$Date, "%Y%m%d ")
```

```{r}
statements_all <- as.vector(statements$Text)
length(statements_all) 
```

The next step was converting the dataset into volatile corpora which is a handful form in the following operations. Below can be seen an example statement before any text preprocessing operations applied.

```{r}
(corpus_all <- VCorpus(VectorSource(statements_all)))
```

```{r}
inspect(corpus_all[[1]])
```

## Preprocessing

We start preprocessing with text cleaning using tm_map() function. We lower each case, remove words from the built-in stopwords list, we remove punctuation, unnecessary whitespaces and numbers. At the end we apply PlainTextDocument() function.

```{r}
corpus_clean <- corpus_all %>% 
    tm_map(tolower) %>%
    tm_map(removeWords, stopwords("en")) %>% 
    tm_map(removePunctuation) %>%
    tm_map(stripWhitespace) %>% 
    tm_map(removeNumbers) %>% 
    tm_map(PlainTextDocument)
```

Below can be seen examples of the statements after above cleaning steps.

```{r}
as.character(corpus_clean[[1]]) 
```

In order to ease operations on the corpus, we modify it into a data frame. 

```{r}
df_corpus <- data.frame(text = unlist(sapply(corpus_clean, `[`, "content")), stringsAsFactors = F)
df_corpus <- df_corpus %>% mutate(doc_id = 1:n())
```

```{r}
df_corpus$text[1]
```

In the next step, we add column with cleaned text to statements data frame. Another important operation is text lemmatization which brings the word to its basic form which is grammatically correct. The example result is printed below.

```{r}
statements_clean <- statements %>% 
  mutate(cleaned_text = df_corpus$text)
```

```{r}
statements_clean$cleaned_text <- lemmatize_strings(statements_clean$cleaned_text)
```

```{r}
statements_clean$cleaned_text[1]
```

At this stage, we also count number of words occuring in the original statement and in the cleaned statement in order to further analyse the trend of length of statements over time.

```{r}
count_cleaned_word <- statements_clean %>%
  unnest_tokens(word_count, cleaned_text) %>%
  count(ID, word_count, sort = T) %>% 
  group_by(ID) %>%
  summarize(word_cleaned_count = sum(n))

statements_clean_count <- left_join(statements_clean, count_cleaned_word, by = 'ID')
```

```{r}
count_word <- statements_clean_count %>%
  unnest_tokens(word_count, Text) %>%
  count(ID, word_count, sort = T) %>% 
  group_by(ID) %>% 
  summarize (word_count = sum(n))

statements_final <- left_join(statements_clean_count, count_word, by = 'ID')
```

## Word counts over time

On the plot below we can see line plots of word counts over time, one for statements before and one after cleaning. The number has increased over time for both types until 2014 when the statements lenght started to decrease. The proportion of number of cleaned words to number of total words is quite constant and oscillates around 60%.

We tried to somehow correlate the changes of length of statements with any event and we discovered that it matches to Fed Chair. Until 2014 the Chairman was Ben Bernanke. After that Janet Yellen took over the role. From February 2018, the Fed Chair is Jerome Powell.

```{r}
myplot <- statements_final %>% 
              select(Date, word_count, word_cleaned_count) %>% 
              ggplot() +
              geom_line(aes(x = Date, 
                            y = word_count), 
                        color = viridis(10)[3]) + 
              geom_line(aes(x = Date, 
                            y = word_cleaned_count), 
                        color = viridis(10)[6]) +
              labs(x = "Date", 
                   y = "Number of words", 
                   title = "Comparison of number of words between original and cleaned <br>statements content over time") +
              scale_x_date(date_breaks = "1 year", 
                           date_labels = "%Y") +
              theme_minimal()

ggplotly(myplot)
```

## Zipf's law

The Zipf's law in context of text mining, states that the frequency of a word is inversely proportional to it's ordered rank. We decided to check the accuracy of the law empirically on our dataset. Below is a dataset with included zipf's frequency.

```{r}
word_counts_zipf <- statements_clean_count %>%
  mutate(year = year(Date)) %>% 
  unnest_tokens(word_count, cleaned_text) %>%
  count(word_count, sort = T)

word_count <- word_counts_zipf 
colnames(word_count) <- c("word", "count")
word_count <- word_count %>%
  mutate(word = factor(word, levels = word),
         rank = row_number(),
         zipfs_freq = ifelse(rank == 1, count, dplyr::first(count) / rank))
```

```{r}
word_count
```

Additionally we decided to present the visualisation on a point plot. Based on this visualisation we decided to remove words with the highest and lowest values, setting the cutoff to ranks 17 and 300 leaving only words in between.

```{r}
p1 <- ggplot(word_count, 
              aes(x = rank, y = count, 
                  color = rank,
                  text = paste("Word: ", word, 
                               "<br>Frequency of word: ", count,
                               "<br>Rank: ", rank))) + 
        geom_point() +
        labs(x = "Rank", y = "count", title = "Zipf's law visualization") +
        scale_color_viridis_c() +
        geom_vline(xintercept = 17) +
        geom_vline(xintercept = 300) +
        theme_minimal() +
        theme(legend.position = "none")
ggplotly(p1, tooltip = "text")
```


```{r}
large_zipf <- as.vector(word_count$word[1:17])
small_zipf <- as.vector(word_count$word[300:1174])
corpus_clean <- corpus_clean %>% tm_map(removeWords, large_zipf)
corpus_clean <- corpus_clean %>% tm_map(removeWords, small_zipf)

df_corpus <- data.frame(text = unlist(sapply(corpus_clean, `[`, "content")), stringsAsFactors = F)
df_corpus <- df_corpus %>% mutate(doc_id = 1:n())

statements_clean <- statements %>% 
  mutate(cleaned_text = df_corpus$text)

statements_clean$cleaned_text <- lemmatize_strings(statements_clean$cleaned_text)
```

## TF-IDF

Using term frequency–inverse document frequency statistic, we proceeded to further analyze the statements to find these words that carry more information than the others. It occured that the top 3 highest TF-IDF value have words *bank*, *central* and *arrangement* in 38 statement. 

```{r}
statements_words <- statements_clean %>%
                      mutate(year = year(Date)) %>% 
                      unnest_tokens(word_count, cleaned_text) %>%
                      count(ID, year, word_count, sort = T)

statements_words_id <- statements_words %>%
                      bind_tf_idf(word_count, ID, n) %>% 
                      arrange(desc(tf_idf))

statements_words_id
```

Using the computed statistic, we visualised the most important words per year. It is interesting to see *balanceconsistent* word in two years (2017 and 2018). Looking at raw data, we can see that it's occurence is caused by the lack of space after dot which was removed by us during preprocessing. It is a clear proof that statements are copy-pasted from previous years even without fixing previous mistakes.

```{r, fig.height=15, fig.width=8, warning=FALSE, message=FALSE}
statements_words_year <- statements_words %>% 
  group_by(year, word_count) %>% 
  summarize(n = sum(n)) %>%  
  bind_tf_idf(word_count, year, n) %>% 
  arrange(desc(tf_idf)) %>% 
  ungroup()

pd <- statements_words_year %>%
      arrange(desc(tf_idf)) %>%
      mutate(word = factor(word_count, levels = rev(unique(word_count)))) %>% 
      group_by(year) %>% 
      top_n(10) %>% 
      ungroup() %>%
      arrange(year, tf_idf) %>%
      mutate(order = row_number()) 

ggplot(pd, aes(order, tf_idf, fill = tf_idf)) +
      geom_bar(show.legend = FALSE, stat = "identity") +
      labs(x = NULL, y = "TF-IDF value") +
      facet_wrap(~year, ncol = 2, scales = "free") +
      scale_x_continuous(breaks = pd$order, labels = pd$word, expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      coord_flip() +
      theme_minimal() +
      scale_fill_viridis_c(direction=-1)
```

## Wordclouds

We also presented the frequency of words using wordclouds. The bigger and lighter the word the higher the frequency. This is much more visual and pleasant to interpret then the table. Based on the cloud below it occured that the most frequent words overall are *securities*, *conditions*, *funds*, *financial* and *range*.

```{r}
dtm <- TermDocumentMatrix(corpus_clean)
m <- as.matrix(dtm)
v <- sort(rowSums(m), decreasing=TRUE)
d <- data.frame(word = names(v), freq=v)
head(d, 10)
```

```{r, warning=FALSE, message=FALSE}
set.seed(1234)
wordcloud2(d %>% arrange(desc(freq)) %>% head(100), color=viridis(100, direction = -1), shape='circle', size=0.2, minRotation = -pi/2, ellipticity = .8, )
```

## Sentiment analysis

Another analysis which we conducted is sentiment extraction for each statement. We decided to use Loughran-McDonald financial dictionary which includes positive and negative categories of sentiment. This dictionary was first presented in the Journal of Finance and has been widely used in the finance domain ever since.

```{r}
tidy_statement <- statements_clean %>%
  mutate(year = year(Date)) %>% 
  group_by(year) %>%
  ungroup() %>%
  unnest_tokens(word, cleaned_text) 

tidy_statement <- tidy_statement %>% 
  select(year, ID, word)
```

```{r}
LM_dict <- loadDictionaryLM()

LM_dict_pos <- as.data.frame(LM_dict$positiveWords)
LM_dict_pos$sentiment <- c("positive")
colnames(LM_dict_pos)[1] <- "word"

LM_dict_neg <- as.data.frame(LM_dict$negativeWords)
LM_dict_neg$sentiment <- c("negative")
colnames(LM_dict_neg)[1] <- "word"

LM_dict <- rbind(LM_dict_pos, LM_dict_neg)
```

Firstly, we check which positive and negative words are the most frequent in FOMC statements. Below can be seen heads of tables listing 5 most frequent positive and negative words.

```{r warning=FALSE, message=FALSE}
statement_word_counts_pos <- tidy_statement %>%
  inner_join(LM_dict) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>% 
  filter(sentiment == "positive")

head(statement_word_counts_pos, 5)
```

```{r warning=FALSE, message=FALSE}
statement_word_counts_neg <- tidy_statement %>%
  inner_join(LM_dict) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>% 
  filter(sentiment == "negative")

head(statement_word_counts_neg, 5)
```

As a final sentiment, we calculate the difference between the sum of positive and negative words in each statement. We presented the results for each statement on the plot below also marking the year. An interesting observation can be noticed from 2008 to 2012 where the sentiment has suddenly turned negative. This can be connected with the 2008 financial crisis. After that we can see a more positive sentiment that was getting higher each year. 

```{r warning=FALSE, message=FALSE}
statement_sentiment <- tidy_statement %>% 
  inner_join(LM_dict) %>% 
  count(year, ID, sentiment) %>% 
  spread(sentiment, n, fill=0) %>% 
  mutate(sentiment = positive - negative)
```

```{r fig.height=10, fig.width=6, warning=FALSE, message=FALSE}
sentiment_plot <- ggplot(statement_sentiment, aes(as.factor(ID), 
                                       sentiment, 
                                       fill = sentiment, 
                                       text = paste("Statement id: ", ID,
                                                    "<br>Sentiment value: ", sentiment))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~year, ncol = 3, scales = "free_x") +
  labs(x="", y="Sentiment") +
  scale_fill_viridis_c(direction=-1) +
  theme_minimal()

ggplotly(sentiment_plot, tooltip = "text") %>% 
  layout(autosize = F, width = 650, height = 800)
```

## Topic modelling

Topic modelling is unsupervised method for identifying topics in documents. LDA is a popular method for topic modelling. In this approach, it is assumed that every document is a combination of few topics and every topic is combination of words. What is important, is fact that one word can be a part of few topics, so topics can be somehow similiar to each other.

In LDA, we have to specify number of topics we want to obtain. In addition to that we can adjust two parameters alpha and beta which correspond to document-topic density and topic-word density.

```{r, warning=FALSE, message=FALSE}
tdm_tm <- DocumentTermMatrix(corpus_clean) 
lda <- LDA(tdm_tm, k = 5, control = list(seed = 1234))
lda_topics <- tidy(lda, matrix = "beta")
```

```{r}
lda_top_terms <- lda_topics %>%
  group_by(topic) %>%
  top_n(8, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

lda_top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE, fill = viridis(40)) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  labs(y="Value of beta parameter", x="") +
  scale_x_reordered() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

In our case we choose 5 topics and based on graphs it is visible that topic number one, three and five can be similar due to the word "securities".

```{r, warning=FALSE, message=FALSE}
by_statement <- statements_clean %>%
  group_by(ID)
 
by_statement_word <- by_statement %>%
  unnest_tokens(word, cleaned_text)

word_counts <- by_statement_word %>%
  anti_join(stop_words) %>%
  count(ID, word, sort = TRUE)

statements_dtm <- word_counts %>%
  cast_dtm(ID, word, n)

statements_lda <- LDA(statements_dtm, k = 5, control = list(seed = 1234))

statements_lda_td <- tidy(statements_lda)

statements_lda_gamma <- tidy(statements_lda, matrix = "gamma")
```



```{r, warning=FALSE, message=FALSE}
topic_1 = statements_lda_gamma %>% filter(., topic=='1')
topic_2 = statements_lda_gamma %>% filter(., topic=='2')
topic_3 = statements_lda_gamma %>% filter(., topic=='3')
topic_4 = statements_lda_gamma %>% filter(., topic=='4')
topic_5 = statements_lda_gamma %>% filter(., topic=='5')

topic_1 = topic_1 %>% rename('topic1' = 'gamma')
topic_2 = topic_2 %>% rename('topic2' = 'gamma')
topic_3 = topic_3 %>% rename('topic3' = 'gamma')
topic_4 = topic_4 %>% rename('topic4' = 'gamma')
topic_5 = topic_5 %>% rename('topic5' = 'gamma')

statements_topics = merge(statements_clean, topic_1, by.x = "ID", by.y = "document")
statements_topics = merge(statements_topics, topic_2, by.x = "ID", by.y = "document")
statements_topics = merge(statements_topics, topic_3, by.x = "ID", by.y = "document")
statements_topics = merge(statements_topics, topic_4, by.x = "ID", by.y = "document")
statements_topics = merge(statements_topics, topic_5, by.x = "ID", by.y = "document")

drops <- c("topic.x", "topic.y")
statements_topics = statements_topics[ , !(names(statements_topics) %in% drops)]
```

It is also possible to see how topics change over time. In our analysis it comes out that every statemnt is always based on 5 topics in similar proportions.

```{r}
colors_viridis <- viridis(5)
topic_density <- ggplot(statements_topics, aes(x=Date)) + 
  geom_line(aes(y=topic1), color=colors_viridis[1]) +
  geom_line(aes(y=topic2), color=colors_viridis[2]) +
  geom_line(aes(y=topic3), color=colors_viridis[3]) +
  geom_line(aes(y=topic4), color=colors_viridis[4]) +
  geom_line(aes(y=topic5), color=colors_viridis[5]) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal()

ggplotly(topic_density)
```


## Word correlation analysis

At the end we were curious which words are mostly correlated. To do so, we create graphs of words, if the words are connected it means that correlation is above given threshold, in our case 65%. It comes out that we have one big graph with nodes that are connected together and few pairs of words.

```{r}
by_statement_word <- data.frame(lapply(by_statement_word, as.character), stringsAsFactors=FALSE)
```

```{r}
# count words co-occuring within sections
word_pairs <- by_statement_word %>%
  pairwise_count(word, ID, sort = TRUE)

word_pairs
```

```{r}
word_cors <- by_statement_word %>%
  group_by(word) %>%
  filter(n() >= 107) %>%
  pairwise_cor(word, ID, sort = TRUE)

word_cors <- word_cors %>% filter(correlation<=1)
```


```{r, warning=FALSE, message=FALSE}
word_cors %>%
  filter(correlation > .65) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_label(aes(label = name), repel = TRUE) +
  theme_void()
```

[^fomc]: Source: https://www.federalreserve.gov/monetarypolicy/fomccalendars.htm.